---
title: "individual1"
author: "Alicia Hauglie"
date: "3/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
```


```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)

offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)

biography <- read_excel('cyclonesFootball2018.xlsx',sheet='Biography')
str(biography)
```

**Part One: Cleaning Data**

1. Player names and opponent names are factors.
```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)
str(defense)

offense$Name <- as.factor(offense$Name)
offense$Opponent_Opponent <- as.factor(offense$Opponent_Opponent)
str(offense)

biography$Name <- as.factor(biography$Name)
str(biography)
```

2. All offensive and defensive statistics are numerical. The Weight column in biogrpahy is also numeric.

```{r}
defense <- mutate_at(defense, vars(3:11), as.numeric)
str(defense)
offense <- mutate_at(offense, vars(3:12), as.numeric)
str(offense)
biography <- mutate_at(biography, vars("Weight"), as.numeric)
str(biography)
```

3. Bonus: Change the Height column in biogrpahy into numeric. (Hint: Use a unit that works well)
```{r}
biography <- biography %>% separate(Height, c("Feet", "Inches"), "-")
biography$Feet <- as.numeric(biography$Feet)
biography$Inches <- as.numeric(biography$Inches)
biography <- biography %>% mutate(Height_in_inches = (Feet*12)+Inches)
biography <- subset(biography, select = -c(Feet, Inches))
str(biography)
```
```{r}
defClean <- defense
offClean <- offense
bioClean <- biography

str(defClean)
str(offClean)
str(bioClean)
```

**Part Two: Tidying (1)**

1. Reformat the defClean data frame into a tidy format using gather. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic.

```{r}
defClean <- defClean %>% gather(key= Statistic, value="Score",3:11)
View(defClean)
```

2. 


**Part Two: Tidying (2)**

1. Separate the Hometown column into two columns, namely the city and the state. Print the first few rows of these two columns.

```{r}
bioClean <- bioClean %>% separate(Hometown, c("City", "State"), ",")
head(bioClean)
```

2. How many players are there from each state?

```{r}
bioCleanstat<- bioClean %>% group_by(State) %>% summarize(n=n())
View(bioCleanstat)
#I don't know if this is right but it looks like something at least lmao
```

**Part Three: Joining Data Frames**

1. Find where do the offensive players come from by joining the offensive data and player biography data appropriately.

```{r}

```

2. Is there a relationship between the weight of a player and the receiving yards (Receiving_YDS)?

```{r}

```

3. Explore more. Raise a question that can be answered only by looking at two or more data frames. Then answer that question.

```{r}

```

