---
title: "team"
author: "Alicia Hauglie & Samuel Cole"
date: "3/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readxl)
```

```{r}
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)

offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)

biography <- read_excel('cyclonesFootball2018.xlsx',sheet='Biography')
str(biography)
```

**Part One: Cleaning Data**

1. Player names and opponent names are factors.
```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)
str(defense)

offense$Name <- as.factor(offense$Name)
offense$Opponent_Opponent <- as.factor(offense$Opponent_Opponent)
str(offense)

biography$Name <- as.factor(biography$Name)
str(biography)
```

2. All offensive and defensive statistics are numerical. The Weight column in biogrpahy is also numeric.

```{r}
defense <- mutate_at(defense, vars(3:11), as.numeric)
str(defense)
offense <- mutate_at(offense, vars(3:12), as.numeric)
str(offense)
biography <- mutate_at(biography, vars("Weight"), as.numeric)
str(biography)
```

3. Bonus: Change the Height column in biogrpahy into numeric. (Hint: Use a unit that works well)
```{r}
biography <- biography %>% separate(Height, c("Feet", "Inches"), "-")
biography$Feet <- as.numeric(biography$Feet)
biography$Inches <- as.numeric(biography$Inches)
biography <- biography %>% mutate(Height_in_inches = (Feet*12)+Inches)
biography <- subset(biography, select = -c(Feet, Inches))
str(biography)
```

```{r}
defClean <- defense
offClean <- offense
bioClean <- biography

str(defClean)
str(offClean)
str(bioClean)
```


**Part Two: Tidying (1)**

1. Reformat the defClean data frame into a tidy format using gather. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic.

```{r}
defClean <- defClean %>% gather(key= Statistic, value="Score",3:11)
```


