---
title: "team"
author: "Alicia Hauglie & Samuel Cole"
date: "3/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)
```

```{r}
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)

offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)

biography <- read_excel('cyclonesFootball2018.xlsx',sheet='Biography')
str(biography)
```

**Part One: Cleaning Data**

1. Player names and opponent names are factors.
```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)
str(defense)

offense$Name <- as.factor(offense$Name)
offense$Opponent_Opponent <- as.factor(offense$Opponent_Opponent)
str(offense)

biography$Name <- as.factor(biography$Name)
str(biography)
```

2. All offensive and defensive statistics are numerical. The Weight column in biogrpahy is also numeric.

```{r}
defense <- mutate_at(defense, vars(3:11), as.numeric)
str(defense)
offense <- mutate_at(offense, vars(3:12), as.numeric)
str(offense)
biography <- mutate_at(biography, vars("Weight"), as.numeric)
str(biography)
```

3. Bonus: Change the Height column in biogrpahy into numeric. (Hint: Use a unit that works well)
```{r}
biography <- biography %>% separate(Height, c("Feet", "Inches"), "-")
biography$Feet <- as.numeric(biography$Feet)
biography$Inches <- as.numeric(biography$Inches)
biography <- biography %>% mutate(Height_in_inches = (Feet*12)+Inches)
biography <- subset(biography, select = -c(Feet, Inches))
str(biography)
```

```{r}
defClean <- defense
offClean <- offense
bioClean <- biography

str(defClean)
str(offClean)
str(bioClean)
```


**Part Two: Tidying (1)**

1. Reformat the defClean data frame into a tidy format using gather. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic.

```{r}
defClean <- defClean %>% gather(key= Statistic, value="Score",3:11)
```

2. Create a histogram for each defensive statistic. Use facet_wrap to draw all histograms in a single plot. What defensive skills are rare?
```{r}
ggplot(data = defClean, aes(x = Score,)) + geom_histogram(binwidth = .5, color="black", fill = "red") + facet_wrap(~Statistic) + ylab("Count of Players in each Score") + xlab("Score") + ggtitle("Histogram of Defensive Statistics")
```
This Graph shows that any type of Turnover is a rare statistic to see. You can also see that Sacks are a rare statistic. Both of these observations make sense with my understanding of football. 


3. Did ISU have better defense against West Virginia or Kansas State? Answer this question by creating a scatterplot with x-axis being the number of solo tackles (by each player) against West Virginia, and y-axis being that against Kansas State. A large number of solo tackles indicates good defense.
```{r}
defClean3 <- defense %>%
  select(Name, Opponent_Opponent, Tackles_Solo) %>%
  spread(Opponent_Opponent, Tackles_Solo)

defClean3 <- defClean3[, c("Kansas State", "West Virginia")]

ggplot(data = defClean3, aes(x = `Kansas State`, y= `West Virginia`)) + geom_point(Na.rm = TRUE) + ylab("Solo Tackles by player against Kansas State") + xlab("Solo Tackles by player against West Virginia") + ggtitle("Comparing ISU's defense against Kansas State and West Virginia") + geom_jitter(Na.rm = TRUE)
```
What we can see is that the defense from Kansas is was better and we had more solo tackles during that game.
**Part Two: Tidying (2)**

1. Separate the Hometown column into two columns, namely the city and the state. Print the first few rows of these two columns.

```{r}
bioClean <- bioClean %>% separate(Hometown, c("City", "State"), ",")
head(bioClean)
```

2. How many players are there from each state?

```{r}
bioCleanstat<- bioClean %>% group_by(State) %>% summarize(n=n())

```

**Part Three: Joining Data Frames**

1. Find where do the offensive players come from by joining the offensive data and player biography data appropriately.

```{r}
origin <- offClean %>% inner_join(bioClean, by='Name')
```

2. Is there a relationship between the weight of a player and the receiving yards (Receiving_YDS)?

```{r}
ggplot(data = origin, aes(x = Weight, y= Receiving_YDS, color = Name)) + geom_point() + ylab("Receiving Yards") + xlab("Weight") + ggtitle("Receiving Yards v.s Player Weight")
```
There is a slight negative relationship between the weight of a player and the recieving yards, meaning that the less a player weighs, the more yards they're likely to have received.


3. Explore more. Raise a question that can be answered only by looking at two or more data frames. Then answer that question.
    Our Question: Which postition has the most Solo tackles during the season?
```{r}
New1 <- defClean %>% inner_join(bioClean, by='Name')
New2 <- New1[New1$Statistic == "Tackles_Solo",]
ggplot(data = New2, aes(x = Position )) + geom_bar( fill = "darkred", color = "black") + coord_flip() 
```
We can see that the Defensive back position has the most tackles by far, and that defensive end and Lineback are 2nd and 3rd in Solo Tackles. The data may be a bit skewed for the tackles for the Defensive back position because our best player on defense was Brian Peavy, and he was a defensive back. 
